---
title: "Healthy Future — Belgium Talk Analysis"
subtitle: "Data analysis for Stanford REAP Conference, Leuven, March 28, 2026"
author: "Sean Sylvia"
date: "2026-03-26"
format:
  html:
    theme: [cosmo, ../style/hf.scss]
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    fig-width: 9
    fig-height: 5.5
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| code-fold: false

library(haven)
library(dplyr)
library(tidyr)
library(purrr)
library(forcats)
library(tibble)
library(stringr)
library(fixest)     # feols — fast, cluster SEs, FEs
library(ggplot2)
library(patchwork)
library(gt)
library(scales)
library(glue)

# Paths
data_path   <- here::here("data", "HF_main_anasam.dta")
fig_path    <- here::here("outputs", "figures")
table_path  <- here::here("outputs", "tables")

dir.create(fig_path,   showWarnings = FALSE, recursive = TRUE)
dir.create(table_path, showWarnings = FALSE, recursive = TRUE)

# Load
df <- read_dta(data_path)
cat(glue("Loaded: {nrow(df)} obs × {ncol(df)} variables\n"))
cat(glue("Treatment: {sum(df$treat==1)} treated, {sum(df$treat==0)} control\n"))

# Brand colors
HF_BLUE   <- "#1f4e79"
HF_TEAL   <- "#2e75b6"
HF_ORANGE <- "#ed7d31"
HF_GREEN  <- "#4a7c59"
HF_GRAY   <- "#767171"

# LASSO-selected controls already in dataset
lasso_cols <- df |> select(starts_with(c("_CHILD_", "_FAM_", "_CARE_", "_CID_"))) |> names()
cat(glue("{length(lasso_cols)} LASSO control variables loaded\n"))

# Helper: star labels
add_stars <- function(p) {
  case_when(p < 0.01 ~ "***", p < 0.05 ~ "**", p < 0.10 ~ "*", TRUE ~ "")
}

# Helper: quote non-syntactic variable names for formulas
bt <- function(x) paste0("`", x, "`")

# Helper: run OLS ITT with cluster SEs via fixest
# Uses pre-selected LASSO controls + county FEs
run_itt <- function(outcome, data = df) {
  controls <- lasso_cols[lasso_cols %in% names(data)]
  rhs <- paste(c("treat", bt(controls)), collapse = " + ")
  fml <- as.formula(paste(bt(outcome), "~", rhs))
  feols(fml, data = data, cluster = ~TID, warn = FALSE)
}
```

---

## Slide 6 — Primary Outcomes

Primary outcomes: hemoglobin level (g/dL), exclusive breastfeeding (0/1), dietary diversity score.

> **Method:** OLS with pre-selected LASSO controls + county FEs, cluster-robust SEs by township (TID). This approximates the `pdslasso` approach used in the Stata paper code, using the `_CHILD_*`, `_FAM_*`, `_CARE_*`, `_CID_*` controls already generated in the dataset.

```{r}
#| label: slide6-primary

primary_vars <- c(
  "FLVAR_CH_HEMO"  = "Hemoglobin Level (g/dL)",
  "FLVAR_FEED_EBF" = "Exclusive Breastfeeding (0/1)",
  "FLVAR_FEED_DDS" = "Dietary Diversity Score"
)

primary_res <- map_dfr(names(primary_vars), function(v) {
  m <- run_itt(v)
  ctrl_mean <- df |> filter(treat == 0) |> pull(v) |> mean(na.rm = TRUE)
  tibble(
    outcome   = primary_vars[[v]],
    b         = coef(m)["treat"],
    se        = se(m)["treat"],
    p         = fixest::pvalue(m)["treat"],
    ci_lo     = coef(m)["treat"] - 1.645 * fixest::se(m)["treat"],
    ci_hi     = coef(m)["treat"] + 1.645 * fixest::se(m)["treat"],
    ctrl_mean = ctrl_mean,
    n         = nobs(m)
  )
})

# Table
primary_res |>
  mutate(
    coeff = glue("{round(b,3)}{add_stars(p)}"),
    se_   = glue("({round(se,3)})"),
    ci    = glue("[{round(ci_lo,3)}, {round(ci_hi,3)}]"),
    p_    = round(p, 4),
    ctrl  = round(ctrl_mean, 3)
  ) |>
  select(Outcome=outcome, Effect=coeff, SE=se_, `90% CI`=ci, `p-value`=p_, `Control Mean`=ctrl, N=n) |>
  gt() |>
  tab_header(title = "Primary Outcomes — ITT Estimates") |>
  tab_footnote("* p<0.10  ** p<0.05  *** p<0.01 | OLS with LASSO controls + county FEs, cluster SEs (township)") |>
  tab_style(style = cell_fill(color = "#e2efda"),
            locations = cells_body(rows = 3)) |>
  gtsave(file.path(table_path, "BT_slide6_primary.html"))
```

```{r}
#| label: slide6-figure
#| fig-cap: "Primary outcomes table — talk-ready"
#| fig-height: 2.2

primary_res |>
  mutate(
    outcome = factor(outcome, levels = rev(outcome)),
    sig     = p < 0.10,
    label   = glue("{round(b,3)}{add_stars(p)}")
  ) |>
  ggplot(aes(y = outcome, x = b, xmin = ci_lo, xmax = ci_hi, color = sig)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = HF_GRAY, linewidth = 0.6) +
  geom_errorbarh(height = 0.2, linewidth = 1) +
  geom_point(aes(shape = sig), size = 4) +
  geom_text(aes(label = label, x = ci_hi), hjust = -0.2, size = 3.5, fontface = "bold") +
  scale_color_manual(values = c("FALSE" = HF_GRAY, "TRUE" = HF_TEAL), guide = "none") +
  scale_shape_manual(values = c("FALSE" = 1, "TRUE" = 19), guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.25))) +
  labs(
    title = "Primary Outcomes",
    x = "Treatment Effect (90% CI)",
    y = NULL,
    caption = "OLS with LASSO controls + county FEs · Cluster SEs by township · Filled = p<0.10"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE),
    plot.caption  = element_text(color = HF_GRAY, size = 8),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide6_primary.png"), width = 8, height = 2.5, dpi = 200)
```

---

## Slide 7 — Secondary Outcomes Forest Plot

```{r}
#| label: slide7-forest
#| fig-cap: "Secondary outcomes — domain index ITT estimates, 90% CIs"
#| fig-height: 5.5

domain_vars <- tribble(
  ~var,           ~label,                        ~group,
  "IN_CH",         "Child Health Index",           "child",
  "IN_BFI",        "Breastfeeding Initiation",     "child",
  "IN_CF",         "Complementary Feeding",        "child",
  "IN_PRIKNOW_CH", "Caregiving Knowledge",         "caregiver",
  "IN_EPDS",       "Maternal Mental Health (EPDS)","caregiver",
  "IN_DASS",       "Caregiver Mental Health (DASS)","caregiver",
  "IN_SS",         "Perceived Social Support",     "caregiver"
)

sec_res <- map_dfr(domain_vars$var, function(v) {
  m <- run_itt(v)
  tibble(
    var   = v,
    b     = coef(m)["treat"],
    se    = se(m)["treat"],
    p     = fixest::pvalue(m)["treat"],
    ci_lo = coef(m)["treat"] - 1.645 * fixest::se(m)["treat"],
    ci_hi = coef(m)["treat"] + 1.645 * fixest::se(m)["treat"],
    n     = nobs(m)
  )
}) |>
  left_join(domain_vars, by = "var") |>
  mutate(
    label = factor(label, levels = rev(domain_vars$label)),
    sig   = p < 0.10,
    coeff_label = glue("{round(b,3)}{add_stars(p)}")
  )

ggplot(sec_res, aes(y = label, x = b, xmin = ci_lo, xmax = ci_hi, color = group)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = HF_GRAY, linewidth = 0.7) +
  geom_errorbarh(height = 0.25, linewidth = 1.1) +
  geom_point(aes(shape = sig), size = 4.5) +
  geom_text(aes(label = coeff_label, x = ci_hi), hjust = -0.15, size = 3.3) +
  # Section headers
  annotate("text", x = -0.28, y = 7.6, label = "Child Outcomes", 
           fontface = "bold", color = HF_BLUE, hjust = 0, size = 3.8) +
  annotate("text", x = -0.28, y = 4.6, label = "Caregiver Outcomes",
           fontface = "bold", color = HF_ORANGE, hjust = 0, size = 3.8) +
  scale_color_manual(
    values = c("child" = HF_BLUE, "caregiver" = HF_ORANGE),
    guide  = "none"
  ) +
  scale_shape_manual(
    values = c("FALSE" = 1, "TRUE" = 19),
    labels = c("p ≥ 0.10", "p < 0.10"),
    name   = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.25))) +
  labs(
    title   = "Treatment Effects — Domain Indices",
    subtitle = "90% confidence intervals · filled = significant at 10%",
    x = "Treatment Effect (SD units)",
    y = NULL,
    caption = "OLS with pre-selected LASSO controls + county FEs · Cluster-robust SEs by township"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 14),
    plot.subtitle = element_text(color = HF_GRAY, size = 10),
    plot.caption  = element_text(color = HF_GRAY, size = 8),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide7_forest.png"), width = 9, height = 5.5, dpi = 200)
```

---

## Slide 8 — Entry Timing Event Study

Effects by child's age at the start of the intervention. Negative age = in utero (enrolled during pregnancy).

```{r}
#| label: slide8-setup

# BBVAR_AGE_MON_BASE2 = age in months at baseline/program start (negative = pregnancy)
df <- df |>
  mutate(
    age_entry = BBVAR_AGE_MON_BASE2,
    entry_bin = case_when(
      treat == 0                       ~ "control",
      age_entry < -3                   ~ "preg_early",   # Early pregnancy
      age_entry >= -3 & age_entry < 0  ~ "preg_late",    # Late pregnancy
      age_entry >= 0  & age_entry < 3  ~ "inf_0_3",      # 0–3m
      age_entry >= 3  & age_entry < 6  ~ "inf_3_6",      # 3–6m
      age_entry >= 6                   ~ "inf_6plus",     # >6m
      TRUE ~ NA_character_
    )
  )

df |> count(entry_bin) |> arrange(entry_bin)
```

```{r}
#| label: slide8-regressions
#| fig-cap: "Entry timing — effects by stage at first exposure, 90% CIs"
#| fig-height: 6.5

event_vars <- tribble(
  ~var,           ~title,
  "IN_CH",         "(a) Child Health Index",
  "IN_DASS",       "(b) Caregiver Mental Health",
  "IN_PRIKNOW_CH", "(c) Caregiving Knowledge",
  "IN_SS",         "(d) Perceived Social Support"
)

bins_order  <- c("preg_early", "preg_late", "inf_0_3", "inf_3_6")
bin_labels  <- c("Preg.\n(early)", "Preg.\n(late)", "Infancy\n0–3m", "Infancy\n3–6m")

# For each outcome × each entry bin: compare that bin vs control
event_res <- map_dfr(event_vars$var, function(v) {
  map_dfr(bins_order, function(b) {
    sub <- df |> filter(entry_bin == b | treat == 0) |>
      mutate(d = as.integer(entry_bin == b & treat == 1))
    controls_here <- lasso_cols[lasso_cols %in% names(sub)]
    fml <- as.formula(paste(bt(v), "~ d +", paste(bt(controls_here), collapse = " + ")))
    m   <- feols(fml, data = sub, cluster = ~TID, warn = FALSE)
    tibble(
      var   = v,
      bin   = b,
      b     = coef(m)["d"],
      se    = se(m)["d"],
      p     = fixest::pvalue(m)["d"],
      ci_lo = coef(m)["d"] - 1.645 * fixest::se(m)["d"],
      ci_hi = coef(m)["d"] + 1.645 * fixest::se(m)["d"]
    )
  })
}) |>
  mutate(
    bin = factor(bin, levels = bins_order, labels = bin_labels),
    sig = p < 0.10
  ) |>
  left_join(event_vars, by = "var") |>
  mutate(title = factor(title, levels = event_vars$title))

# 2×2 panel
ggplot(event_res, aes(x = bin, y = b, ymin = ci_lo, ymax = ci_hi, color = sig, group = 1)) +
  # Pregnancy shading
  annotate("rect", xmin = 0.5, xmax = 2.5, ymin = -Inf, ymax = Inf,
           fill = "#dce6f1", alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.6) +
  geom_line(color = HF_BLUE, linewidth = 0.7, alpha = 0.5) +
  geom_errorbar(width = 0.18, linewidth = 1) +
  geom_point(aes(shape = sig), size = 3.5) +
  # Pregnancy / Infancy label
  geom_vline(xintercept = 2.5, linetype = "dotted", color = HF_GRAY, linewidth = 0.7) +
  annotate("text", x = 1.5, y = Inf, label = "Pregnancy", vjust = 1.5,
           size = 3.2, color = HF_GRAY, fontface = "italic") +
  annotate("text", x = 3.5, y = Inf, label = "Infancy",   vjust = 1.5,
           size = 3.2, color = HF_GRAY, fontface = "italic") +
  scale_color_manual(values = c("FALSE" = HF_GRAY, "TRUE" = HF_TEAL), guide = "none") +
  scale_shape_manual(values = c("FALSE" = 1, "TRUE" = 19), guide = "none") +
  coord_cartesian(ylim = c(-0.5, 0.65)) +
  facet_wrap(~title, ncol = 2) +
  labs(
    title    = "Effects by Stage at First Exposure",
    subtitle = "90% confidence intervals · each group vs. control",
    x = "Entry timing",
    y = "Treatment effect (SD units)",
    caption  = "OLS with LASSO controls + county FEs · Cluster SEs by township"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 13),
    plot.subtitle = element_text(color = HF_GRAY, size = 9),
    plot.caption  = element_text(color = HF_GRAY, size = 8),
    strip.text    = element_text(face = "bold", size = 10, color = HF_BLUE),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide8_event.png"), width = 9, height = 6.5, dpi = 200)
```

---

## Slide 9 — Intermediate Outcomes Predict Child Health

**New analysis.** Control group only. Standardized associations between intermediate outcomes at follow-up and the child health index — motivating the surrogate index framework for HF 2.0.

```{r}
#| label: slide9-predictive
#| fig-cap: "Intermediate outcomes predicting child health (control group)"
#| fig-height: 4.5

ctrl_df <- df |> filter(treat == 0)
cat(glue("Control group N = {nrow(ctrl_df)}\n"))

# Standardize
z_score <- function(x) (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)

surrogates <- tribble(
  ~var,           ~label,
  "IN_CF",         "Complementary Feeding",
  "IN_BFI",        "Breastfeeding Initiation",
  "IN_PRIKNOW_CH", "Caregiving Knowledge",
  "IN_DASS",       "Caregiver Mental Health",
  "IN_SS",         "Perceived Social Support"
)

ctrl_std <- ctrl_df |>
  mutate(
    across(c(all_of(surrogates$var), "IN_CH"), z_score, .names = "z_{.col}")
  )

ctrl_county <- lasso_cols[lasso_cols %in% names(ctrl_std)]

surr_res <- map_dfr(surrogates$var, function(v) {
  fml <- as.formula(paste(bt("z_IN_CH"), "~", bt(paste0("z_", v)), "+", paste(bt(ctrl_county), collapse = " + ")))
  m   <- feols(fml, data = ctrl_std, cluster = ~TID, warn = FALSE)
  coef_name <- paste0("z_", v)
  tibble(
    var   = v,
    b     = coef(m)[coef_name],
    se    = se(m)[coef_name],
    p     = fixest::pvalue(m)[coef_name],
    ci_lo = coef(m)[coef_name] - 1.96 * fixest::se(m)[coef_name],
    ci_hi = coef(m)[coef_name] + 1.96 * fixest::se(m)[coef_name],
    n     = nobs(m)
  )
}) |>
  left_join(surrogates, by = "var") |>
  mutate(
    label = factor(label, levels = rev(surrogates$label)),
    sig   = p < 0.10,
    coeff_label = glue("{round(b,3)}{add_stars(p)}")
  )

# Save table
surr_res |>
  mutate(
    b_   = glue("{round(b,3)}{add_stars(p)}"),
    se_  = glue("({round(se,3)})"),
    ci_  = glue("[{round(ci_lo,3)}, {round(ci_hi,3)}]"),
    p_   = round(p, 4)
  ) |>
  select(Outcome=label, Coeff=b_, SE=se_, `95% CI`=ci_, `p-value`=p_, N=n) |>
  gt() |>
  tab_header(
    title    = "Intermediate Outcomes Predict Child Health",
    subtitle = "Control group only · Standardized coefficients"
  ) |>
  tab_footnote("All variables in SD units. OLS with county FEs. Cluster SEs by township.") |>
  gtsave(file.path(table_path, "BT_slide9_predictive.html"))

# Figure
ggplot(surr_res, aes(y = label, x = b, xmin = ci_lo, xmax = ci_hi, color = sig)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = HF_GRAY, linewidth = 0.7) +
  geom_errorbarh(height = 0.25, linewidth = 1.2) +
  geom_point(aes(shape = sig), size = 5) +
  geom_text(aes(label = coeff_label, x = ci_hi), hjust = -0.15, size = 3.8, fontface = "bold") +
  scale_color_manual(values = c("FALSE" = HF_GRAY, "TRUE" = HF_ORANGE), guide = "none") +
  scale_shape_manual(values = c("FALSE" = 1, "TRUE" = 19), guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.30))) +
  labs(
    title    = "Intermediate Outcomes Predict Child Health Trajectories",
    subtitle = "Control group only · 95% CI · all variables standardized (SD units)",
    x        = "Association with Child Health Index (SD units)",
    y        = NULL,
    caption  = glue(
      "OLS with county FEs · Cluster-robust SEs by township · N ≈ {min(surr_res$n)}\n",
      "Families with better intermediate outcomes at follow-up show stronger child health. ",
      "Motivates surrogate index framework for HF 2.0."
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 13),
    plot.subtitle = element_text(color = HF_GRAY, size = 10),
    plot.caption  = element_text(color = HF_GRAY, size = 8.5),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide9_predictive.png"), width = 9, height = 4.5, dpi = 200)
```

---

## Output Summary

```{r}
#| label: summary
#| code-fold: false

figs   <- list.files(fig_path,   pattern = "*.png", full.names = TRUE)
tables <- list.files(table_path, pattern = "*.html|*.csv", full.names = TRUE)

cat("=== Figures ===\n")
walk(figs, ~cat(sprintf("  %-45s  %s KB\n", basename(.x), round(file.size(.x)/1024))))

cat("\n=== Tables ===\n")
walk(tables, ~cat(sprintf("  %s\n", basename(.x))))
```
