---
title: "Healthy Future — Belgium Talk Analysis"
subtitle: "Data analysis for Stanford REAP Conference, Leuven, March 28, 2026"
author: "Sean Sylvia"
date: "2026-03-26"
format:
  html:
    theme: [cosmo, ../style/hf.scss]
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    fig-width: 9
    fig-height: 5.5
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| code-fold: false

library(haven)
library(dplyr)
library(tidyr)
library(purrr)
library(forcats)
library(tibble)
library(stringr)
library(fixest)     # feols — fast, cluster SEs, FEs
library(ggplot2)
library(patchwork)
library(gt)
library(scales)
library(glue)

# Paths
data_path   <- here::here("data", "HF_main_anasam.dta")
fig_path    <- here::here("outputs", "figures")
table_path  <- here::here("outputs", "tables")

dir.create(fig_path,   showWarnings = FALSE, recursive = TRUE)
dir.create(table_path, showWarnings = FALSE, recursive = TRUE)

# Load
df <- read_dta(data_path)
cat(glue("Loaded: {nrow(df)} obs × {ncol(df)} variables\n"))
cat(glue("Treatment: {sum(df$treat==1)} treated, {sum(df$treat==0)} control\n"))

# Brand colors
HF_BLUE   <- "#1f4e79"
HF_TEAL   <- "#2e75b6"
HF_ORANGE <- "#ed7d31"
HF_GREEN  <- "#4a7c59"
HF_GRAY   <- "#767171"

# LASSO-selected controls already in dataset
lasso_cols <- df |> select(starts_with(c("_CHILD_", "_FAM_", "_CARE_", "_CID_"))) |> names()
cat(glue("{length(lasso_cols)} LASSO control variables loaded\n"))

# Helper: star labels
add_stars <- function(p) {
  case_when(p < 0.01 ~ "***", p < 0.05 ~ "**", p < 0.10 ~ "*", TRUE ~ "")
}

# Helper: quote non-syntactic variable names for formulas
bt <- function(x) paste0("`", x, "`")

# Helper: run OLS ITT with cluster SEs via fixest
# Uses pre-selected LASSO controls + county FEs
run_itt <- function(outcome, data = df) {
  controls <- lasso_cols[lasso_cols %in% names(data)]
  rhs <- paste(c("treat", bt(controls)), collapse = " + ")
  fml <- as.formula(paste(bt(outcome), "~", rhs))
  feols(fml, data = data, cluster = ~TID, warn = FALSE)
}
```

---

## Slide 6 — Primary Outcomes

Primary outcomes: hemoglobin level (g/dL), exclusive breastfeeding (0/1), dietary diversity score.

> **Method:** OLS with pre-selected LASSO controls + county FEs, cluster-robust SEs by township (TID). This approximates the `pdslasso` approach used in the Stata paper code, using the `_CHILD_*`, `_FAM_*`, `_CARE_*`, `_CID_*` controls already generated in the dataset.

```{r}
#| label: slide6-primary

primary_vars <- c(
  "FLVAR_CH_HEMO"  = "Hemoglobin Level (g/dL)",
  "FLVAR_FEED_EBF" = "Exclusive Breastfeeding (0/1)",
  "FLVAR_FEED_DDS" = "Dietary Diversity Score"
)

primary_res <- map_dfr(names(primary_vars), function(v) {
  m <- run_itt(v)
  ctrl_mean <- df |> filter(treat == 0) |> pull(v) |> mean(na.rm = TRUE)
  tibble(
    outcome   = primary_vars[[v]],
    b         = coef(m)["treat"],
    se        = se(m)["treat"],
    p         = fixest::pvalue(m)["treat"],
    ci_lo     = coef(m)["treat"] - 1.645 * fixest::se(m)["treat"],
    ci_hi     = coef(m)["treat"] + 1.645 * fixest::se(m)["treat"],
    ctrl_mean = ctrl_mean,
    n         = nobs(m)
  )
})

# Table
primary_res |>
  mutate(
    coeff = glue("{round(b,3)}{add_stars(p)}"),
    se_   = glue("({round(se,3)})"),
    ci    = glue("[{round(ci_lo,3)}, {round(ci_hi,3)}]"),
    p_    = round(p, 4),
    ctrl  = round(ctrl_mean, 3)
  ) |>
  select(Outcome=outcome, Effect=coeff, SE=se_, `90% CI`=ci, `p-value`=p_, `Control Mean`=ctrl, N=n) |>
  gt() |>
  tab_header(title = "Primary Outcomes — ITT Estimates") |>
  tab_footnote("* p<0.10  ** p<0.05  *** p<0.01 | OLS with LASSO controls + county FEs, cluster SEs (township)") |>
  tab_style(style = cell_fill(color = "#e2efda"),
            locations = cells_body(rows = 3)) |>
  gtsave(file.path(table_path, "BT_slide6_primary.html"))
```

```{r}
#| label: slide6-figure
#| fig-cap: "Primary outcomes table — talk-ready"
#| fig-height: 2.2

primary_res |>
  mutate(
    outcome = factor(outcome, levels = rev(outcome)),
    sig     = p < 0.10,
    label   = glue("{round(b,3)}{add_stars(p)}")
  ) |>
  ggplot(aes(y = outcome, x = b, xmin = ci_lo, xmax = ci_hi, color = sig)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = HF_GRAY, linewidth = 0.6) +
  geom_errorbarh(height = 0.2, linewidth = 1) +
  geom_point(aes(shape = sig), size = 4) +
  geom_text(aes(label = label, x = ci_hi), hjust = -0.2, size = 3.5, fontface = "bold") +
  scale_color_manual(values = c("FALSE" = HF_GRAY, "TRUE" = HF_TEAL), guide = "none") +
  scale_shape_manual(values = c("FALSE" = 1, "TRUE" = 19), guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.25))) +
  labs(
    title = "Primary Outcomes",
    x = "Treatment Effect (90% CI)",
    y = NULL,
    caption = "OLS with LASSO controls + county FEs · Cluster SEs by township · Filled = p<0.10"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE),
    plot.caption  = element_text(color = HF_GRAY, size = 8),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide6_primary.png"), width = 8, height = 2.5, dpi = 200)
```

---

## Slide 7 — Secondary Outcomes Forest Plot

```{r}
#| label: slide7-forest
#| fig-cap: "Secondary outcomes — domain index ITT estimates, 90% CIs"
#| fig-height: 5.5

domain_vars <- tribble(
  ~var,           ~label,                        ~group,
  "IN_CH",         "Child Health Index",           "child",
  "IN_BFI",        "Breastfeeding Initiation",     "child",
  "IN_CF",         "Complementary Feeding",        "child",
  "IN_PRIKNOW_CH", "Caregiving Knowledge",         "caregiver",
  "IN_EPDS",       "Maternal Mental Health (EPDS)","caregiver",
  "IN_DASS",       "Caregiver Mental Health (DASS)","caregiver",
  "IN_SS",         "Perceived Social Support",     "caregiver"
)

sec_res <- map_dfr(domain_vars$var, function(v) {
  m <- run_itt(v)
  tibble(
    var   = v,
    b     = coef(m)["treat"],
    se    = se(m)["treat"],
    p     = fixest::pvalue(m)["treat"],
    ci_lo = coef(m)["treat"] - 1.645 * fixest::se(m)["treat"],
    ci_hi = coef(m)["treat"] + 1.645 * fixest::se(m)["treat"],
    n     = nobs(m)
  )
}) |>
  left_join(domain_vars, by = "var") |>
  mutate(
    label = factor(label, levels = rev(domain_vars$label)),
    sig   = p < 0.10,
    coeff_label = glue("{round(b,3)}{add_stars(p)}")
  )

ggplot(sec_res, aes(y = label, x = b, xmin = ci_lo, xmax = ci_hi, color = group)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = HF_GRAY, linewidth = 0.7) +
  geom_errorbarh(height = 0.25, linewidth = 1.1) +
  geom_point(aes(shape = sig), size = 4.5) +
  geom_text(aes(label = coeff_label, x = ci_hi), hjust = -0.15, size = 3.3) +
  # Section headers
  annotate("text", x = -0.28, y = 7.6, label = "Child Outcomes", 
           fontface = "bold", color = HF_BLUE, hjust = 0, size = 3.8) +
  annotate("text", x = -0.28, y = 4.6, label = "Caregiver Outcomes",
           fontface = "bold", color = HF_ORANGE, hjust = 0, size = 3.8) +
  scale_color_manual(
    values = c("child" = HF_BLUE, "caregiver" = HF_ORANGE),
    guide  = "none"
  ) +
  scale_shape_manual(
    values = c("FALSE" = 1, "TRUE" = 19),
    labels = c("p ≥ 0.10", "p < 0.10"),
    name   = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.25))) +
  labs(
    title   = "Treatment Effects — Domain Indices",
    subtitle = "90% confidence intervals · filled = significant at 10%",
    x = "Treatment Effect (SD units)",
    y = NULL,
    caption = "OLS with pre-selected LASSO controls + county FEs · Cluster-robust SEs by township"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 14),
    plot.subtitle = element_text(color = HF_GRAY, size = 10),
    plot.caption  = element_text(color = HF_GRAY, size = 8),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide7_forest.png"), width = 9, height = 5.5, dpi = 200)
```

---

## Slide 8 — Entry Timing Event Study

Effects by child's age at the start of the intervention. Negative age = in utero (enrolled during pregnancy).

```{r}
#| label: slide8-setup

# BBVAR_AGE_MON_BASE2 = age in months at baseline/program start (negative = pregnancy)
df <- df |>
  mutate(
    age_entry = BBVAR_AGE_MON_BASE2,
    entry_bin = case_when(
      treat == 0                       ~ "control",
      age_entry < -3                   ~ "preg_early",   # Early pregnancy
      age_entry >= -3 & age_entry < 0  ~ "preg_late",    # Late pregnancy
      age_entry >= 0  & age_entry < 3  ~ "inf_0_3",      # 0–3m
      age_entry >= 3  & age_entry < 6  ~ "inf_3_6",      # 3–6m
      age_entry >= 6                   ~ "inf_6plus",     # >6m
      TRUE ~ NA_character_
    )
  )

df |> count(entry_bin) |> arrange(entry_bin)
```

```{r}
#| label: slide8-regressions
#| fig-cap: "Entry timing — effects by stage at first exposure, 90% CIs"
#| fig-height: 6.5

event_vars <- tribble(
  ~var,           ~title,
  "IN_CH",         "(a) Child Health Index",
  "IN_DASS",       "(b) Caregiver Mental Health",
  "IN_PRIKNOW_CH", "(c) Caregiving Knowledge",
  "IN_SS",         "(d) Perceived Social Support"
)

bins_order  <- c("preg_early", "preg_late", "inf_0_3", "inf_3_6")
bin_labels  <- c("Preg.\n(early)", "Preg.\n(late)", "Infancy\n0–3m", "Infancy\n3–6m")

# For each outcome × each entry bin: compare that bin vs control
event_res <- map_dfr(event_vars$var, function(v) {
  map_dfr(bins_order, function(b) {
    sub <- df |> filter(entry_bin == b | treat == 0) |>
      mutate(d = as.integer(entry_bin == b & treat == 1))
    controls_here <- lasso_cols[lasso_cols %in% names(sub)]
    fml <- as.formula(paste(bt(v), "~ d +", paste(bt(controls_here), collapse = " + ")))
    m   <- feols(fml, data = sub, cluster = ~TID, warn = FALSE)
    tibble(
      var   = v,
      bin   = b,
      b     = coef(m)["d"],
      se    = se(m)["d"],
      p     = fixest::pvalue(m)["d"],
      ci_lo = coef(m)["d"] - 1.645 * fixest::se(m)["d"],
      ci_hi = coef(m)["d"] + 1.645 * fixest::se(m)["d"]
    )
  })
}) |>
  mutate(
    bin = factor(bin, levels = bins_order, labels = bin_labels),
    sig = p < 0.10
  ) |>
  left_join(event_vars, by = "var") |>
  mutate(title = factor(title, levels = event_vars$title))

# 2×2 panel
ggplot(event_res, aes(x = bin, y = b, ymin = ci_lo, ymax = ci_hi, color = sig, group = 1)) +
  # Pregnancy shading
  annotate("rect", xmin = 0.5, xmax = 2.5, ymin = -Inf, ymax = Inf,
           fill = "#dce6f1", alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.6) +
  geom_line(color = HF_BLUE, linewidth = 0.7, alpha = 0.5) +
  geom_errorbar(width = 0.18, linewidth = 1) +
  geom_point(aes(shape = sig), size = 3.5) +
  # Pregnancy / Infancy label
  geom_vline(xintercept = 2.5, linetype = "dotted", color = HF_GRAY, linewidth = 0.7) +
  annotate("text", x = 1.5, y = Inf, label = "Pregnancy", vjust = 1.5,
           size = 3.2, color = HF_GRAY, fontface = "italic") +
  annotate("text", x = 3.5, y = Inf, label = "Infancy",   vjust = 1.5,
           size = 3.2, color = HF_GRAY, fontface = "italic") +
  scale_color_manual(values = c("FALSE" = HF_GRAY, "TRUE" = HF_TEAL), guide = "none") +
  scale_shape_manual(values = c("FALSE" = 1, "TRUE" = 19), guide = "none") +
  coord_cartesian(ylim = c(-0.5, 0.65)) +
  facet_wrap(~title, ncol = 2) +
  labs(
    title    = "Effects by Stage at First Exposure",
    subtitle = "90% confidence intervals · each group vs. control",
    x = "Entry timing",
    y = "Treatment effect (SD units)",
    caption  = "OLS with LASSO controls + county FEs · Cluster SEs by township"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 13),
    plot.subtitle = element_text(color = HF_GRAY, size = 9),
    plot.caption  = element_text(color = HF_GRAY, size = 8),
    strip.text    = element_text(face = "bold", size = 10, color = HF_BLUE),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide8_event.png"), width = 9, height = 6.5, dpi = 200)
```

---

## Slide 9 — Intermediate Outcomes Predict Child Health

**New analysis.** Control group only. Standardized associations between intermediate outcomes at follow-up and the child health index — motivating the surrogate index framework for HF 2.0.

```{r}
#| label: slide9-predictive
#| fig-cap: "Intermediate outcomes predicting child health (control group)"
#| fig-height: 4.5

ctrl_df <- df |> filter(treat == 0)
cat(glue("Control group N = {nrow(ctrl_df)}\n"))

# Standardize
z_score <- function(x) (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)

surrogates <- tribble(
  ~var,           ~label,
  "IN_CF",         "Complementary Feeding",
  "IN_BFI",        "Breastfeeding Initiation",
  "IN_PRIKNOW_CH", "Caregiving Knowledge",
  "IN_DASS",       "Caregiver Mental Health",
  "IN_SS",         "Perceived Social Support"
)

ctrl_std <- ctrl_df |>
  mutate(
    across(c(all_of(surrogates$var), "IN_CH"), z_score, .names = "z_{.col}")
  )

ctrl_county <- lasso_cols[lasso_cols %in% names(ctrl_std)]

surr_res <- map_dfr(surrogates$var, function(v) {
  fml <- as.formula(paste(bt("z_IN_CH"), "~", bt(paste0("z_", v)), "+", paste(bt(ctrl_county), collapse = " + ")))
  m   <- feols(fml, data = ctrl_std, cluster = ~TID, warn = FALSE)
  coef_name <- paste0("z_", v)
  tibble(
    var   = v,
    b     = coef(m)[coef_name],
    se    = se(m)[coef_name],
    p     = fixest::pvalue(m)[coef_name],
    ci_lo = coef(m)[coef_name] - 1.96 * fixest::se(m)[coef_name],
    ci_hi = coef(m)[coef_name] + 1.96 * fixest::se(m)[coef_name],
    n     = nobs(m)
  )
}) |>
  left_join(surrogates, by = "var") |>
  mutate(
    label = factor(label, levels = rev(surrogates$label)),
    sig   = p < 0.10,
    coeff_label = glue("{round(b,3)}{add_stars(p)}")
  )

# Save table
surr_res |>
  mutate(
    b_   = glue("{round(b,3)}{add_stars(p)}"),
    se_  = glue("({round(se,3)})"),
    ci_  = glue("[{round(ci_lo,3)}, {round(ci_hi,3)}]"),
    p_   = round(p, 4)
  ) |>
  select(Outcome=label, Coeff=b_, SE=se_, `95% CI`=ci_, `p-value`=p_, N=n) |>
  gt() |>
  tab_header(
    title    = "Intermediate Outcomes Predict Child Health",
    subtitle = "Control group only · Standardized coefficients"
  ) |>
  tab_footnote("All variables in SD units. OLS with county FEs. Cluster SEs by township.") |>
  gtsave(file.path(table_path, "BT_slide9_predictive.html"))

# Figure
ggplot(surr_res, aes(y = label, x = b, xmin = ci_lo, xmax = ci_hi, color = sig)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = HF_GRAY, linewidth = 0.7) +
  geom_errorbarh(height = 0.25, linewidth = 1.2) +
  geom_point(aes(shape = sig), size = 5) +
  geom_text(aes(label = coeff_label, x = ci_hi), hjust = -0.15, size = 3.8, fontface = "bold") +
  scale_color_manual(values = c("FALSE" = HF_GRAY, "TRUE" = HF_ORANGE), guide = "none") +
  scale_shape_manual(values = c("FALSE" = 1, "TRUE" = 19), guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.30))) +
  labs(
    title    = "Intermediate Outcomes Predict Child Health Trajectories",
    subtitle = "Control group only · 95% CI · all variables standardized (SD units)",
    x        = "Association with Child Health Index (SD units)",
    y        = NULL,
    caption  = glue(
      "OLS with county FEs · Cluster-robust SEs by township · N ≈ {min(surr_res$n)}\n",
      "Families with better intermediate outcomes at follow-up show stronger child health. ",
      "Motivates surrogate index framework for HF 2.0."
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 13),
    plot.subtitle = element_text(color = HF_GRAY, size = 10),
    plot.caption  = element_text(color = HF_GRAY, size = 8.5),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide9_predictive.png"), width = 9, height = 4.5, dpi = 200)
```

---

## Slide 10 — CATE Estimation with Causal Forests {#slide10}

> **Research question:** Does the Healthy Future program benefit all families equally, or are there important subgroups who respond more (or less) to the intervention?

```{r}
#| label: slide10-cate-setup
#| code-fold: true

library(grf)

# Covariates: raw baseline variables (not LASSO-transformed controls)
cate_vars <- c(
  "BBVAR_AGE_MON", "BBVAR_BOY",
  "HHVAR_MOM_AGE_BASE", "HHVAR_MOM_HIEDU", "HHVAR_ASSET",
  "HHVAR_SIB", "HHVAR_SIZE_BASE", "HHVAR_MOM_MIGHIS", "HHVAR_MOM_HOMETOWN",
  "BLVAR_DASSPRI_DEP", "BLVAR_DASSPRI_ANX", "BLVAR_DASSPRI_STR",
  "BLVAR_KNOWPRI_INX", "BLVAR_SSPRI_ALL", "BLVAR_MFEED_DDS", "BLVAR_FEED_EBF"
)

# Complete cases on outcome + treatment + covariates + cluster
cf_df <- df |>
  select(all_of(c("FLVAR_FEED_DDS", "treat", "TID", cate_vars))) |>
  drop_na()

cat(glue("Causal forest sample: {nrow(cf_df)} obs (dropped {nrow(df) - nrow(cf_df)} with missing data)\n"))

Y        <- as.numeric(cf_df$FLVAR_FEED_DDS)
W        <- as.numeric(cf_df$treat)
X        <- cf_df |> select(all_of(cate_vars)) |> as.matrix()
clusters <- as.integer(as.factor(cf_df$TID))

# Fit causal forest
set.seed(2026)
cf <- causal_forest(X, Y, W,
                    clusters  = clusters,
                    num.trees = 2000,
                    seed      = 2026)

# ATE from forest
ate <- average_treatment_effect(cf)

cat(glue(
  "\n=== ATE from Causal Forest ===\n",
  "  Estimate : {round(ate['estimate'], 4)}\n",
  "  Std. Err.: {round(ate['std.err'], 4)}\n",
  "  95% CI   : [{round(ate['estimate'] - 1.96*ate['std.err'], 4)},",
  " {round(ate['estimate'] + 1.96*ate['std.err'], 4)}]\n"
))
```

### CATE Distribution

```{r}
#| label: slide10-cate-distribution
#| fig-cap: "Distribution of estimated conditional average treatment effects (CATEs) for dietary diversity. Vertical line = ATE from the forest."

# Predicted CATEs
cates <- predict(cf)$predictions

ate_est <- ate[["estimate"]]
ate_se  <- ate[["std.err"]]

cate_df <- tibble(cate = cates)

ggplot(cate_df, aes(x = cate)) +
  geom_histogram(binwidth = 0.05, fill = HF_TEAL, color = "white", alpha = 0.85) +
  geom_vline(xintercept = ate_est, color = HF_ORANGE, linewidth = 1.4, linetype = "solid") +
  geom_vline(xintercept = ate_est - 1.96 * ate_se, color = HF_ORANGE,
             linewidth = 0.7, linetype = "dashed") +
  geom_vline(xintercept = ate_est + 1.96 * ate_se, color = HF_ORANGE,
             linewidth = 0.7, linetype = "dashed") +
  annotate("text", x = ate_est + 0.02, y = Inf, label = glue("ATE = {round(ate_est, 3)}"),
           vjust = 2, hjust = 0, color = HF_ORANGE, fontface = "bold", size = 4) +
  labs(
    title    = "Distribution of Estimated CATEs \u2014 Dietary Diversity",
    subtitle = glue("Causal forest · N = {nrow(cf_df)} · 2,000 trees · clustered by township"),
    x        = "Estimated CATE (additional food groups)",
    y        = "Count",
    caption  = "Wager & Athey (2018). Orange line = ATE; dashed = 95% CI."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 13),
    plot.subtitle = element_text(color = HF_GRAY, size = 10),
    plot.caption  = element_text(color = HF_GRAY, size = 8.5),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide10_cate_distribution.png"), width = 9, height = 5, dpi = 200)
```

### Variable Importance

```{r}
#| label: slide10-var-importance
#| fig-cap: "Variable importance from causal forest — which baseline characteristics most drive treatment effect heterogeneity?"

vi <- variable_importance(cf)
vi_df <- tibble(
  variable   = cate_vars,
  importance = as.numeric(vi)
) |>
  arrange(desc(importance)) |>
  slice_head(n = 10) |>
  mutate(variable = fct_reorder(variable, importance))

ggplot(vi_df, aes(x = importance, y = variable)) +
  geom_col(fill = HF_BLUE, alpha = 0.85) +
  geom_text(aes(label = round(importance, 3)), hjust = -0.1, size = 3.5, color = HF_GRAY) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(
    title    = "Which Baseline Characteristics Drive Treatment Effect Heterogeneity?",
    subtitle = "Variable importance from causal forest (higher = more predictive of CATE variation)",
    x        = "Importance",
    y        = NULL,
    caption  = "Variable importance = mean decrease in MSE from splits on each variable."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 13),
    plot.subtitle = element_text(color = HF_GRAY, size = 10),
    plot.caption  = element_text(color = HF_GRAY, size = 8.5),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide10_var_importance.png"), width = 9, height = 5, dpi = 200)
```

### Best Linear Projection (BLP) Test for Heterogeneity

```{r}
#| label: slide10-blp
#| code-fold: false

blp <- test_calibration(cf)
print(blp)

# coeftest returns a matrix with cols: Estimate, Std. Error, t value, Pr(>|t|)
blp_mat  <- as.matrix(blp)
blp_coef <- blp_mat["differential.forest.prediction", "Estimate"]
blp_se   <- blp_mat["differential.forest.prediction", "Std. Error"]
blp_pval <- blp_mat["differential.forest.prediction", ncol(blp_mat)]  # last col = p-value

cat(glue(
  "\n=== BLP Test for Heterogeneity ===\n",
  "  Differential forest prediction: {round(blp_coef, 4)}\n",
  "  Std. Err.                      : {round(blp_se, 4)}\n",
  "  p-value                        : {round(blp_pval, 4)}\n",
  "\nInterpretation: A coefficient > 1 (or p < 0.05) indicates meaningful heterogeneity",
  " — families predicted to benefit more actually do benefit more.\n"
))
```

### Top-K Targeting Analysis

```{r}
#| label: slide10-topk
#| fig-cap: "Average treatment effects for families sorted by predicted CATE. Top 20% = highest predicted responders."

cate_rank_df <- cf_df |>
  mutate(
    cate  = cates,
    Y     = Y,
    W     = W
  )

# Compute ATE within each targeting group using IPW-style split
n_total <- nrow(cate_rank_df)
thresh_20 <- quantile(cate_rank_df$cate, 0.80)
thresh_50 <- quantile(cate_rank_df$cate, 0.50)

compute_group_ate <- function(sub_df) {
  # Simple Neyman estimator
  mu1 <- mean(sub_df$Y[sub_df$W == 1])
  mu0 <- mean(sub_df$Y[sub_df$W == 0])
  n1  <- sum(sub_df$W == 1)
  n0  <- sum(sub_df$W == 0)
  se  <- sqrt(var(sub_df$Y[sub_df$W == 1]) / n1 + var(sub_df$Y[sub_df$W == 0]) / n0)
  tibble(est = mu1 - mu0, se = se, n = nrow(sub_df))
}

groups <- list(
  "Top 20%\n(highest predicted)"   = cate_rank_df |> filter(cate >= thresh_20),
  "Top 50%\n(above median)"        = cate_rank_df |> filter(cate >= thresh_50),
  "Bottom 50%\n(below median)"     = cate_rank_df |> filter(cate <  thresh_50),
  "All families\n(overall ATE)"    = cate_rank_df
)

topk_res <- map_dfr(groups, compute_group_ate, .id = "group") |>
  mutate(
    ci_lo = est - 1.96 * se,
    ci_hi = est + 1.96 * se,
    sig   = ci_lo > 0 | ci_hi < 0,
    group = fct_inorder(group)
  )

ggplot(topk_res, aes(x = group, y = est, fill = sig)) +
  geom_col(alpha = 0.85, width = 0.6) +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi), width = 0.2, linewidth = 1.1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = HF_GRAY) +
  geom_text(aes(label = glue("{round(est, 3)}\n(n={n})"),
                y = ci_hi + 0.01), size = 3.5, vjust = 0, color = HF_BLUE, fontface = "bold") +
  scale_fill_manual(values = c("FALSE" = HF_GRAY, "TRUE" = HF_TEAL), guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.20))) +
  labs(
    title    = "Treatment Effects by Predicted Benefit Group",
    subtitle = "Families ranked by predicted CATE \u2014 Neyman estimator \u00b7 95% CI",
    x        = NULL,
    y        = "ATE (additional food groups)",
    caption  = "Top-K analysis: sorted by causal forest CATE predictions."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 13),
    plot.subtitle = element_text(color = HF_GRAY, size = 10),
    plot.caption  = element_text(color = HF_GRAY, size = 8.5),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide10_topk_targeting.png"), width = 9, height = 5, dpi = 200)
```

---

## Slide 11 — RATE Analysis: Gains from Personalization {#slide11}

> **Research question:** How much do we gain by targeting the intervention toward families predicted to benefit most, compared to treating everyone?

```{r}
#| label: slide11-rate
#| code-fold: true

# Compute RATE (Rank-Weighted Average Treatment Effect)
rate <- rank_average_treatment_effect(cf, cf$predictions)

# Extract RATE estimate and SE
rate_est  <- rate$estimate
rate_se   <- rate$std.err
rate_pval <- 2 * pnorm(-abs(rate_est / rate_se))

cat(glue(
  "=== RATE Analysis ===\n",
  "  RATE estimate: {round(rate_est, 4)}\n",
  "  Std. Err.    : {round(rate_se, 4)}\n",
  "  95% CI       : [{round(rate_est - 1.96*rate_se, 4)}, {round(rate_est + 1.96*rate_se, 4)}]\n",
  "  p-value      : {round(rate_pval, 4)}\n"
))
```

### Targeting Operator Characteristic (TOC) Curve

```{r}
#| label: slide11-toc-curve
#| fig-cap: "TOC curve showing treatment effect gains from targeting. The shaded area equals the RATE — the gain from prioritizing predicted high-responders."

# Build TOC curve from rate object
# Inspect actual column names
toc_raw <- as_tibble(rate$TOC)
# Column names may vary; standardize
names(toc_raw) <- tolower(names(toc_raw))
# Expected: something like "fraction", "estimate", "std.err" or "se"
toc_df <- toc_raw |>
  rename_with(~ case_when(
    . %in% c("fraction", "q")          ~ "fraction",
    . %in% c("estimate", "toc")        ~ "toc_est",
    . %in% c("std.err", "se", "stderr")~ "toc_se",
    TRUE                               ~ .
  )) |>
  mutate(
    ci_lo = toc_est - 1.96 * toc_se,
    ci_hi = toc_est + 1.96 * toc_se
  )

ate_line <- ate[["estimate"]]

ggplot(toc_df, aes(x = fraction, y = toc_est)) +
  # Shaded area between TOC curve and ATE line (= RATE gain)
  geom_ribbon(aes(ymin = pmin(toc_est, ate_line),
                  ymax = pmax(toc_est, ate_line)),
              fill = HF_TEAL, alpha = 0.25) +
  # 95% CI ribbon
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = HF_BLUE, alpha = 0.15) +
  # TOC curve
  geom_line(color = HF_BLUE, linewidth = 1.4) +
  # ATE horizontal reference
  geom_hline(yintercept = ate_line, color = HF_ORANGE, linewidth = 1.1, linetype = "dashed") +
  annotate("text", x = 0.05, y = ate_line + 0.01, label = glue("ATE = {round(ate_line, 3)}"),
           color = HF_ORANGE, fontface = "bold", size = 3.8, hjust = 0) +
  # RATE annotation
  annotate("text", x = 0.5, y = max(toc_df$toc_est, na.rm = TRUE) * 0.85,
           label = glue("RATE = {round(rate_est, 3)}\n(p = {round(rate_pval, 3)})"),
           color = HF_BLUE, fontface = "bold", size = 4, hjust = 0.5) +
  scale_x_continuous(labels = scales::percent, breaks = seq(0, 1, 0.2)) +
  labs(
    title    = "Targeting Operator Characteristic \u2014 Gains from Personalization",
    subtitle = glue(
      "RATE = {round(rate_est, 3)} (SE = {round(rate_se, 3)}, p = {round(rate_pval, 3)}) \u00b7 ",
      "Shaded area = gain over uniform treatment"
    ),
    x        = "Fraction of population treated (ranked by predicted benefit)",
    y        = "Average treatment effect for treated fraction",
    caption  = paste0(
      "Yadlowsky et al. (2021) RATE. ",
      "X-axis: 0 = treat nobody; 1 = treat everyone. ",
      "Curve above ATE line = targeting gains."
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 13),
    plot.subtitle = element_text(color = HF_GRAY, size = 10),
    plot.caption  = element_text(color = HF_GRAY, size = 8.5),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_slide11_toc_curve.png"), width = 9, height = 5.5, dpi = 200)
```

### RATE Interpretation

```{r}
#| label: slide11-rate-interp
#| code-fold: false

# Fraction of ATE captured by top-20% targeting
top20_ate <- topk_res |> filter(str_starts(group, "Top 20")) |> pull(est)
all_ate   <- topk_res |> filter(str_starts(group, "All")) |> pull(est)
pct_captured <- round(100 * top20_ate / all_ate, 1)

cat(glue(
  "=== Targeting Interpretation ===\n",
  "  Overall ATE              : {round(all_ate, 4)} food groups\n",
  "  ATE for top 20%          : {round(top20_ate, 4)} food groups\n",
  "  % of avg effect captured : {pct_captured}%\n",
  "\n",
  "If we could target the top 20% of predicted high-responders,\n",
  "we would achieve {round(top20_ate, 3)} food groups improvement vs.\n",
  "{round(all_ate, 3)} for the overall population — capturing {pct_captured}%\n",
  "of the per-family benefit at 20% of the program cost.\n",
  "\n",
  "RATE = {round(rate_est, 4)} (p = {round(rate_pval, 4)})\n",
  if (rate_pval < 0.05) {
    "=> Statistically significant targeting gain: heterogeneity is actionable.\n"
  } else {
    "=> Targeting gain not statistically significant at 5% level.\n"
  }
))
```

---

## CATE Heterogeneity Screening Across All Outcomes {#cate-screen}

> **Research question:** For which outcomes does the Healthy Future program show meaningful treatment effect heterogeneity? We screen all primary and secondary outcomes using causal forests + Best Linear Projection (BLP) tests to identify where personalization might matter most.

```{r}
#| label: cate-screen-setup
#| code-fold: true

# Use the SAME covariate set as the dietary diversity CATE (Slide 10)
# cate_vars is already defined above

# All outcome variables to screen
outcomes_to_screen <- tribble(
  ~variable,             ~label,                              ~domain,
  "FLVAR_CH_HEMO",       "Hemoglobin Level (g/dL)",           "Primary",
  "FLVAR_FEED_EBF",      "Exclusive Breastfeeding",           "Primary",
  "FLVAR_FEED_DDS",      "Dietary Diversity Score",           "Primary",
  "IN_CH",               "Child Health Index",                "Secondary",
  "IN_BFI",              "Breastfeeding Initiation Index",    "Secondary",
  "IN_CF",               "Complementary Feeding Index",       "Secondary",
  "IN_PRIKNOW_CH",       "Caregiving Knowledge Index",        "Secondary",
  "IN_EPDS",             "Maternal Mental Health (EPDS)",     "Secondary",
  "IN_DASS",             "Caregiver Mental Health (DASS)",    "Secondary",
  "IN_SS",               "Perceived Social Support Index",    "Secondary",
  "FLVAR_EFEED_EIBF",    "Early Initiation of BF",           "Breastfeeding",
  "FLVAR_EFEED_COLOS",   "Colostrum Feeding",                "Breastfeeding",
  "FLVAR_EFEED_EvBF",    "Ever Breastfed",                   "Breastfeeding",
  "FLVAR_FEED_MDD",      "Minimum Dietary Diversity",         "Feeding",
  "FLVAR_FEED_MAD",      "Minimum Acceptable Diet",           "Feeding"
)

# Build base complete-case dataset on covariates + treatment + cluster
# Outcomes are allowed to have NAs (handled per-outcome below)
outcome_vars_present <- outcomes_to_screen$variable[
  outcomes_to_screen$variable %in% names(df)
]

base_df <- df |>
  select(all_of(c("treat", "TID", cate_vars)),
         any_of(outcome_vars_present)) |>
  filter(complete.cases(across(all_of(c("treat", "TID", cate_vars)))))

cat(glue("Base covariate-complete sample: {nrow(base_df)} obs (of {nrow(df)} total)\n"))
cat(glue("Outcome variables present in data: {length(outcome_vars_present)} of {nrow(outcomes_to_screen)}\n"))

# Pre-compute covariate matrix, treatment, and cluster vectors for the base sample
X_base       <- base_df |> select(all_of(cate_vars)) |> as.matrix()
W_base       <- as.numeric(base_df$treat)
cluster_base <- as.integer(as.factor(base_df$TID))

# Screening function: fit causal_forest and return BLP test results
run_cate_screen <- function(outcome_var, outcome_label, df_complete, X_mat, W_vec, cluster_vec) {
  if (!outcome_var %in% names(df_complete)) {
    message(glue("  SKIP: {outcome_var} not found in data"))
    return(NULL)
  }
  
  Y <- as.numeric(df_complete[[outcome_var]])
  if (sum(!is.na(Y)) < 50) {
    message(glue("  SKIP: {outcome_var} has fewer than 50 non-missing observations"))
    return(NULL)
  }
  
  # Subset to non-missing outcome rows
  ok <- !is.na(Y)
  
  tryCatch({
    cf_s <- causal_forest(
      X_mat[ok, , drop = FALSE],
      Y[ok],
      W_vec[ok],
      clusters  = cluster_vec[ok],
      num.trees = 2000,
      seed      = 2026
    )
    
    ate_s <- average_treatment_effect(cf_s)
    blp_s <- test_calibration(cf_s)
    
    # test_calibration returns a coeftest matrix
    # Row 2 = "differential.forest.prediction" 
    blp_mat <- as.matrix(blp_s)
    
    # Safely extract differential prediction row
    diff_row <- if ("differential.forest.prediction" %in% rownames(blp_mat)) {
      blp_mat["differential.forest.prediction", ]
    } else {
      blp_mat[2, ]
    }
    
    # p-value is last column
    blp_p_val <- diff_row[length(diff_row)]
    
    tibble(
      outcome         = outcome_label,
      variable        = outcome_var,
      n               = sum(ok),
      ate_est         = ate_s[["estimate"]],
      ate_se          = ate_s[["std.err"]],
      blp_coef        = diff_row[["Estimate"]],
      blp_se          = diff_row[["Std. Error"]],
      blp_p           = blp_p_val,
      has_heterogeneity = blp_p_val < 0.10
    )
  }, error = function(e) {
    message(glue("  ERROR in {outcome_var}: {e$message}"))
    NULL
  })
}
```

### Run Causal Forests for All Outcomes

```{r}
#| label: cate-screen-run
#| cache: false

# Loop over all outcomes present in the data
cat("Running causal forests for all outcomes...\n")
cat("(This may take a few minutes — 2,000 trees per outcome)\n\n")

screen_results <- map2_dfr(
  outcomes_to_screen$variable,
  outcomes_to_screen$label,
  function(v, lbl) {
    cat(glue("  -> {lbl} ({v})\n"))
    run_cate_screen(
      outcome_var   = v,
      outcome_label = lbl,
      df_complete   = base_df,
      X_mat         = X_base,
      W_vec         = W_base,
      cluster_vec   = cluster_base
    )
  }
)

# Join domain information
screen_results <- screen_results |>
  left_join(outcomes_to_screen |> select(variable, domain), by = "variable") |>
  mutate(
    domain = factor(domain, levels = c("Primary", "Secondary", "Breastfeeding", "Feeding")),
    outcome = factor(outcome, levels = outcome)  # preserve order
  )

cat(glue("\nCompleted: {nrow(screen_results)} outcomes screened\n"))
cat(glue("Significant heterogeneity (BLP p < 0.10): {sum(screen_results$has_heterogeneity)} outcomes\n"))
print(screen_results |> filter(has_heterogeneity) |> select(outcome, variable, blp_coef, blp_p))
```

### Summary Table

```{r}
#| label: cate-screen-table

screen_tbl_data <- screen_results |>
  arrange(domain, blp_p) |>
  mutate(
    ate_str  = glue("{round(ate_est, 3)} ({round(ate_se, 3)})"),
    blp_str  = glue("{round(blp_coef, 3)} ({round(blp_se, 3)})"),
    blp_p_   = round(blp_p, 4),
    flag     = if_else(has_heterogeneity, "✓", ""),
    het_flag = has_heterogeneity  # keep for styling
  )

# Row indices with heterogeneity (after arrange)
sig_rows <- which(screen_tbl_data$het_flag)

screen_tbl_data |>
  select(
    Domain = domain,
    Outcome = outcome,
    N = n,
    `ATE (SE)` = ate_str,
    `BLP Coef. (SE)` = blp_str,
    `BLP p-value` = blp_p_,
    `Heterogeneity` = flag
  ) |>
  gt(groupname_col = "Domain") |>
  tab_header(
    title    = "Heterogeneity Screening — BLP Test Across All Outcomes",
    subtitle = "Causal forest + Best Linear Projection test · 2,000 trees · clustered by township"
  ) |>
  tab_footnote(
    footnote = "BLP test (Chernozhukov et al. 2018): coefficient on differential forest prediction. p < 0.10 suggests meaningful heterogeneity. ATE from forest using out-of-bag predictions.",
    locations = cells_title()
  ) |>
  tab_style(
    style     = cell_fill(color = "#e2efda"),
    locations = cells_body(rows = sig_rows)
  ) |>
  tab_style(
    style     = list(cell_text(weight = "bold")),
    locations = cells_body(columns = `BLP p-value`, rows = sig_rows)
  ) |>
  cols_align(align = "center", columns = c(N, `BLP p-value`, Heterogeneity)) |>
  gtsave(file.path(table_path, "BT_cate_screen_blp.html"))
```

### BLP p-value Plot

```{r}
#| label: cate-screen-figure
#| fig-cap: "BLP test p-values across all outcomes. Points left of the dashed line (p = 0.10) indicate meaningful treatment effect heterogeneity."
#| fig-height: 6.5

screen_results |>
  mutate(
    outcome = fct_reorder(outcome, blp_p),
    sig_label = if_else(has_heterogeneity, "p < 0.10 (heterogeneity)", "p ≥ 0.10 (homogeneous)")
  ) |>
  ggplot(aes(x = blp_p, y = outcome, color = has_heterogeneity, shape = domain)) +
  geom_vline(xintercept = 0.10, linetype = "dashed", color = HF_ORANGE, linewidth = 1.0) +
  geom_point(size = 4.5, alpha = 0.9) +
  geom_text(
    data = ~ filter(., has_heterogeneity),
    aes(label = glue("p={round(blp_p,3)}")),
    hjust = -0.15, size = 3.2, fontface = "bold"
  ) +
  annotate("text", x = 0.10, y = 0.5,
           label = "p = 0.10", color = HF_ORANGE,
           hjust = -0.1, vjust = 0, size = 3.5, fontface = "italic") +
  scale_color_manual(
    values = c("FALSE" = HF_GRAY, "TRUE" = HF_GREEN),
    labels = c("p ≥ 0.10", "p < 0.10 ✓"),
    name   = "Heterogeneity"
  ) +
  scale_shape_manual(
    values = c("Primary" = 19, "Secondary" = 17, "Breastfeeding" = 15, "Feeding" = 18),
    name   = "Domain"
  ) +
  scale_x_continuous(
    limits = c(0, NA),
    expand = expansion(mult = c(0.02, 0.25)),
    breaks = c(0, 0.10, 0.25, 0.50, 0.75, 1.00)
  ) +
  labs(
    title    = "Which Outcomes Show Meaningful Treatment Effect Heterogeneity?",
    subtitle = "BLP test (Best Linear Projection) · Causal forest · 2,000 trees",
    x        = "BLP p-value (differential forest prediction)",
    y        = NULL,
    caption  = "Chernozhukov et al. (2018). Dashed line = p = 0.10 threshold. Green = significant heterogeneity."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", color = HF_BLUE, size = 13),
    plot.subtitle = element_text(color = HF_GRAY, size = 10),
    plot.caption  = element_text(color = HF_GRAY, size = 8.5),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_path, "BT_cate_screen_blp.png"), width = 9, height = 6.5, dpi = 200)
```

### Interpretation

```{r}
#| label: cate-screen-interp
#| code-fold: false

# Summarize findings
n_sig <- sum(screen_results$has_heterogeneity)
sig_outcomes <- screen_results |>
  filter(has_heterogeneity) |>
  arrange(blp_p) |>
  pull(outcome)

if (n_sig > 0) {
  cat(glue(
    "=== Heterogeneity Screening Results ===\n\n",
    "Outcomes with significant heterogeneity (BLP p < 0.10):\n"
  ))
  walk2(sig_outcomes, 
        screen_results |> filter(has_heterogeneity) |> arrange(blp_p) |> pull(blp_p),
        ~cat(glue("  - {.x}: BLP p = {round(.y, 4)}\n")))
  cat(glue(
    "\n{n_sig} of {nrow(screen_results)} outcomes show meaningful heterogeneity.\n",
    "These are the strongest candidates for full RATE analysis and targeting.\n"
  ))
} else {
  cat(glue(
    "No outcomes showed significant heterogeneity at p < 0.10.\n",
    "Total outcomes screened: {nrow(screen_results)}\n",
    "This suggests treatment effects are broadly consistent across the population.\n"
  ))
}

# Recommendation
cat(glue(
  "\n=== Recommended for Full RATE Analysis ===\n",
  "Outcomes with BLP p < 0.20 (broadened threshold):\n"
))
screen_results |>
  filter(blp_p < 0.20) |>
  arrange(blp_p) |>
  select(outcome, blp_p, ate_est, ate_se) |>
  mutate(across(where(is.numeric), ~round(., 4))) |>
  print()
```

::: {.callout-note}
**Heterogeneity Screening — Key Takeaways**

The BLP test systematically assessed whether the Healthy Future program's treatment effects vary meaningfully across families for each of the 15 screened outcomes. Outcomes with a significant BLP p-value (< 0.10) indicate that the causal forest's predictions of *who benefits more* are statistically validated — the families predicted to benefit most actually do benefit most when evaluated out-of-bag.

`r if(sum(screen_results$has_heterogeneity) > 0) { glue("**{sum(screen_results$has_heterogeneity)} outcome(s) show significant heterogeneity:** {paste(screen_results$outcome[screen_results$has_heterogeneity], collapse=', ')}. These are the strongest candidates for a full RATE (Rank-Weighted Average Treatment Effect) analysis and precision targeting in HF 2.0.") } else { "No outcomes showed statistically significant heterogeneity at the 10% level, suggesting that HF's treatment effects are broadly homogeneous across the population studied. This is itself informative: the program may benefit families similarly regardless of baseline characteristics, which reduces the case for selective targeting but strengthens the case for broad scale-up." }`

For HF 2.0 design, outcomes exhibiting heterogeneity suggest that baseline assessments (child age, maternal mental health, household assets, baseline feeding practices) could be used to identify which families would benefit most from intensified or tailored intervention components — the core value proposition of a learning health system.
:::

---

## Output Summary

```{r}
#| label: summary
#| code-fold: false

figs   <- list.files(fig_path,   pattern = "*.png", full.names = TRUE)
tables <- list.files(table_path, pattern = "*.html|*.csv", full.names = TRUE)

cat("=== Figures ===\n")
walk(figs, ~cat(sprintf("  %-45s  %s KB\n", basename(.x), round(file.size(.x)/1024))))

cat("\n=== Tables ===\n")
walk(tables, ~cat(sprintf("  %s\n", basename(.x))))
```
