---
title: "Treatment Effect Heterogeneity — Domain Summary Indices"
subtitle: "Healthy Future RCT · Stanford REAP Conference Analysis"
author: "Sean Sylvia"
date: "2026-03-27"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    fig-width: 10
    fig-height: 6
    mainfont: "Source Sans Pro"
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
#| code-fold: false

library(haven)
library(dplyr)
library(tidyr)
library(grf)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(glue)
library(forcats)

# Brand colors
HF_BLUE   <- "#1f4e79"
HF_TEAL   <- "#2e75b6"
HF_ORANGE <- "#ed7d31"
HF_GREEN  <- "#4a7c59"
HF_GRAY   <- "#767171"

# CATE covariates (15 vars, BLVAR_FEED_EBF excluded)
cate_vars <- c(
  "BBVAR_AGE_MON", "BBVAR_BOY",
  "HHVAR_MOM_AGE_BASE", "HHVAR_MOM_HIEDU", "HHVAR_ASSET",
  "HHVAR_SIB", "HHVAR_SIZE_BASE", "HHVAR_MOM_MIGHIS", "HHVAR_MOM_HOMETOWN",
  "BLVAR_DASSPRI_DEP", "BLVAR_DASSPRI_ANX", "BLVAR_DASSPRI_STR",
  "BLVAR_KNOWPRI_INX", "BLVAR_SSPRI_ALL", "BLVAR_MFEED_DDS"
)

cate_labels <- c(
  "Child age (months)", "Boy", "Mother's age", "Mother high education",
  "Asset index", "Number of siblings", "Household size",
  "Mother migration history", "Mother hometown",
  "Baseline depression", "Baseline anxiety", "Baseline stress",
  "Baseline knowledge", "Baseline social support", "Baseline dietary diversity"
)

add_stars <- function(p) {
  case_when(p < 0.01 ~ "***", p < 0.05 ~ "**", p < 0.10 ~ "*", TRUE ~ "")
}

# Load data with indices
df <- readRDS(here::here("data", "hf_with_indices.rds"))
cat(glue("Dataset: {nrow(df)} obs · {sum(df$treat==1)} treated · {sum(df$treat==0)} control\n"))
```

# Data & Index Construction

Three domain summary indices are constructed using Anderson (2008) inverse-covariance-weighted (ICW) aggregation, based on the 3-factor structure identified via exploratory factor analysis.

```{r indices-table}
#| fig-cap: "Anderson ICW Domain Indices"

domains <- list(
  list(name = "idx_mental",  label = "Mental Health",     components = c("IN_EPDS", "IN_DASS"),
       comp_labels = c("Maternal MH (EPDS)", "Caregiver MH (DASS)")),
  list(name = "idx_support", label = "Caregiver Support", components = c("IN_SS", "IN_PRIKNOW_CH"),
       comp_labels = c("Social Support", "Caregiving Knowledge")),
  list(name = "idx_child",   label = "Child Nutrition",   components = c("IN_CH", "IN_CF", "IN_BFI"),
       comp_labels = c("Child Health", "Complementary Feeding", "BF Initiation"))
)

# Descriptive statistics table
desc_rows <- purrr::map_dfr(domains, function(d) {
  x <- df[[d$name]]
  tibble(
    Index       = d$label,
    Components  = paste(d$comp_labels, collapse = " + "),
    N           = sum(!is.na(x)),
    `Ctrl Mean` = round(mean(x[df$treat == 0], na.rm = TRUE), 3),
    `Ctrl SD`   = round(sd(x[df$treat == 0], na.rm = TRUE), 3),
    `Treat Mean`= round(mean(x[df$treat == 1], na.rm = TRUE), 3),
    `Diff`      = round(mean(x[df$treat == 1], na.rm = TRUE) - mean(x[df$treat == 0], na.rm = TRUE), 3)
  )
})

desc_rows |>
  kable(align = "llccccc") |>
  kable_styling(font_size = 13, bootstrap_options = c("striped", "hover")) |>
  row_spec(0, bold = TRUE, background = HF_BLUE, color = "white") |>
  footnote(general = "Anderson (2008) ICW indices. Standardized to control-group mean = 0, SD = 1.")
```

---

# CATE Estimation, BLP, & RATE

```{r cate-all}
#| code-fold: true

# Run causal forest for each domain index
results <- list()

for (d in domains) {
  cat(glue("\n--- {d$label} ({d$name}) ---\n"))

  cf_df <- df |>
    select(all_of(c(d$name, "treat", "TID", cate_vars))) |>
    drop_na()

  Y <- as.numeric(cf_df[[d$name]])
  W <- as.numeric(cf_df$treat)
  X <- cf_df |> select(all_of(cate_vars)) |> as.matrix()
  clusters <- as.integer(as.factor(cf_df$TID))

  set.seed(2026)
  cf <- causal_forest(X, Y, W,
                      clusters = clusters,
                      num.trees = 4000,
                      tune.parameters = "all",
                      seed = 2026)

  ate   <- average_treatment_effect(cf)
  blp   <- test_calibration(cf)
  blp_m <- as.matrix(blp)
  blp_p <- blp_m["differential.forest.prediction", ncol(blp_m)]

  cates <- predict(cf)$predictions
  rate  <- rank_average_treatment_effect(cf, cates)
  rate_p <- 2 * pnorm(-abs(rate$estimate / rate$std.err))

  vi <- variable_importance(cf)

  results[[d$name]] <- list(
    domain = d,
    n      = nrow(cf_df),
    ate_est = ate[["estimate"]], ate_se = ate[["std.err"]],
    blp_coef = blp_m["differential.forest.prediction", "Estimate"],
    blp_se   = blp_m["differential.forest.prediction", "Std. Error"],
    blp_p    = blp_p,
    rate_est = rate$estimate, rate_se = rate$std.err, rate_p = rate_p,
    cate_sd  = sd(cates),
    cates    = cates,
    cf       = cf,
    cf_df    = cf_df,
    vi       = setNames(as.numeric(vi), cate_vars)
  )

  cat(glue("  N={nrow(cf_df)}, ATE={round(ate[['estimate']],4)}, BLP p={round(blp_p,4)}, RATE={round(rate$estimate,4)} (p={round(rate_p,4)})\n"))
}
```

## Summary Table

```{r summary-table}
summary_df <- purrr::map_dfr(results, function(r) {
  tibble(
    Index     = r$domain$label,
    N         = r$n,
    `ATE`     = sprintf("%.3f%s", r$ate_est, add_stars(2*pnorm(-abs(r$ate_est/r$ate_se)))),
    `(SE)`    = sprintf("(%.3f)", r$ate_se),
    `BLP`     = sprintf("%.3f", r$blp_coef),
    `BLP SE`  = sprintf("(%.3f)", r$blp_se),
    `BLP p`   = sprintf("%.3f", r$blp_p),
    `RATE`    = sprintf("%.3f", r$rate_est),
    `RATE SE` = sprintf("(%.3f)", r$rate_se),
    `RATE p`  = sprintf("%.3f", r$rate_p),
    `CATE SD` = sprintf("%.3f", r$cate_sd)
  )
})

summary_df |>
  kable(align = "lcccccccccc") |>
  kable_styling(font_size = 12, bootstrap_options = c("striped", "hover")) |>
  row_spec(0, bold = TRUE, background = HF_BLUE, color = "white") |>
  footnote(
    general = "Causal forest: 4,000 trees, auto-tuned, 15 covariates, clustered by township.",
    symbol  = c("BLP: Best Linear Projection test for heterogeneity (Chernozhukov et al. 2018)",
                "RATE: Rank-Average Treatment Effect (Yadlowsky et al. 2021)",
                "* p<0.10, ** p<0.05, *** p<0.01")
  )
```

---

# Four-Panel Figures

```{r figure-function}
#| code-fold: true

make_four_panel <- function(r, save_path = NULL) {
  cates   <- r$cates
  ate_est <- r$ate_est
  label   <- r$domain$label

  # 1. CATE Distribution
  p1 <- ggplot(tibble(cate = cates), aes(x = cate)) +
    geom_histogram(aes(y = after_stat(density)), bins = 30,
                   fill = HF_TEAL, color = "white", alpha = 0.8) +
    geom_density(color = HF_BLUE, linewidth = 1) +
    geom_vline(xintercept = ate_est, color = HF_ORANGE, linewidth = 1.3) +
    annotate("text", x = ate_est, y = Inf, label = glue("ATE = {round(ate_est, 3)}"),
             vjust = 2, hjust = -0.1, color = HF_ORANGE, fontface = "bold", size = 3.8) +
    labs(title = "(a) CATE Distribution", x = "Estimated CATE", y = "Density") +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold", color = HF_BLUE))

  # 2. Variable Importance
  vi_df <- tibble(variable = cate_labels, importance = r$vi) |>
    arrange(desc(importance)) |> slice_head(n = 8) |>
    mutate(variable = fct_reorder(variable, importance))

  p2 <- ggplot(vi_df, aes(x = importance, y = variable)) +
    geom_col(fill = HF_BLUE, alpha = 0.85) +
    geom_text(aes(label = round(importance, 3)), hjust = -0.1, size = 3, color = HF_GRAY) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.25))) +
    labs(title = "(b) Variable Importance", x = "Importance", y = NULL) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold", color = HF_BLUE))

  # 3. GATES (quintile ATEs)
  cf_df <- r$cf_df
  cf_df$cate <- cates
  cf_df$quintile <- cut(cates, breaks = quantile(cates, probs = 0:5/5),
                        include.lowest = TRUE, labels = paste0("Q", 1:5))

  gates <- cf_df |>
    group_by(quintile) |>
    summarise(
      mu1 = mean(.data[[r$domain$name]][treat == 1], na.rm = TRUE),
      mu0 = mean(.data[[r$domain$name]][treat == 0], na.rm = TRUE),
      n1  = sum(treat == 1),
      n0  = sum(treat == 0),
      .groups = "drop"
    ) |>
    mutate(
      ate = mu1 - mu0,
      se  = sqrt((mu1*(1-mu1)/pmax(n1,1)) + (mu0*(1-mu0)/pmax(n0,1))),
      # Neyman SE for continuous outcomes
      ate_se = NA_real_
    )

  # Recompute SE properly for continuous outcomes
  gates_proper <- cf_df |>
    group_by(quintile) |>
    summarise(
      ate = mean(.data[[r$domain$name]][treat == 1], na.rm = TRUE) -
            mean(.data[[r$domain$name]][treat == 0], na.rm = TRUE),
      se  = sqrt(var(.data[[r$domain$name]][treat == 1], na.rm = TRUE) / sum(treat == 1) +
                 var(.data[[r$domain$name]][treat == 0], na.rm = TRUE) / sum(treat == 0)),
      n   = n(),
      .groups = "drop"
    )

  p3 <- ggplot(gates_proper, aes(x = quintile, y = ate)) +
    geom_hline(yintercept = ate_est, linetype = "dashed", color = HF_ORANGE, linewidth = 0.8) +
    geom_col(fill = HF_TEAL, alpha = 0.7, width = 0.6) +
    geom_errorbar(aes(ymin = ate - 1.96*se, ymax = ate + 1.96*se),
                  width = 0.2, linewidth = 0.8, color = HF_BLUE) +
    geom_text(aes(label = round(ate, 3), y = ate + sign(ate)*1.96*se),
              vjust = ifelse(gates_proper$ate >= 0, -0.5, 1.5), size = 3, fontface = "bold") +
    labs(title = "(c) GATES — Quintile Treatment Effects",
         subtitle = "Q1 = lowest predicted · Q5 = highest predicted",
         x = "CATE Quintile", y = "Group ATE (95% CI)") +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold", color = HF_BLUE),
          plot.subtitle = element_text(color = HF_GRAY, size = 9))

  # 4. TOC Curve
  rate_obj <- rank_average_treatment_effect(r$cf, cates)
  toc_raw  <- as_tibble(rate_obj$TOC)
  names(toc_raw) <- tolower(names(toc_raw))
  toc <- toc_raw |>
    rename_with(~case_when(
      . %in% c("fraction","q")           ~ "fraction",
      . %in% c("estimate","toc")         ~ "toc_est",
      . %in% c("std.err","se","stderr")  ~ "toc_se",
      TRUE ~ .
    )) |>
    mutate(ci_lo = toc_est - 1.96*toc_se, ci_hi = toc_est + 1.96*toc_se)

  p4 <- ggplot(toc, aes(x = fraction, y = toc_est)) +
    geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = HF_BLUE, alpha = 0.12) +
    geom_line(color = HF_BLUE, linewidth = 1.2) +
    geom_hline(yintercept = ate_est, color = HF_ORANGE, linewidth = 0.8, linetype = "dashed") +
    annotate("text", x = 0.05, y = ate_est, label = glue("ATE = {round(ate_est, 3)}"),
             vjust = -1, hjust = 0, color = HF_ORANGE, fontface = "bold", size = 3.5) +
    annotate("text", x = 0.5, y = max(toc$toc_est, na.rm = TRUE) * 0.9,
             label = glue("RATE = {round(r$rate_est, 3)} (p = {round(r$rate_p, 3)})"),
             color = HF_BLUE, fontface = "bold", size = 3.8) +
    scale_x_continuous(labels = scales::percent) +
    labs(title = "(d) TOC Curve — Targeting Gains",
         x = "Fraction treated (ranked by predicted benefit)",
         y = "ATE for treated fraction") +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold", color = HF_BLUE))

  # Combine
  combined <- (p1 | p2) / (p3 | p4) +
    plot_annotation(
      title = glue("Treatment Effect Heterogeneity: {label}"),
      subtitle = glue("N = {r$n} · ATE = {round(ate_est, 3)} · BLP p = {round(r$blp_p, 3)} · RATE = {round(r$rate_est, 3)} (p = {round(r$rate_p, 3)})"),
      theme = theme(
        plot.title = element_text(face = "bold", color = HF_BLUE, size = 14),
        plot.subtitle = element_text(color = HF_GRAY, size = 11)
      )
    )

  if (!is.null(save_path)) {
    ggsave(save_path, combined, width = 14, height = 10, dpi = 200)
    cat(glue("  Saved: {basename(save_path)}\n"))
  }

  combined
}
```

## Mental Health Index

```{r fig-mental}
#| fig-height: 10
#| fig-width: 14
#| fig-cap: "Treatment effect heterogeneity — Mental Health Index (EPDS + DASS)"

make_four_panel(results[["idx_mental"]],
                save_path = here::here("outputs", "figures", "BT_het_mental.png"))
```

## Caregiver Support Index

```{r fig-support}
#| fig-height: 10
#| fig-width: 14
#| fig-cap: "Treatment effect heterogeneity — Caregiver Support Index (SS + Knowledge)"

make_four_panel(results[["idx_support"]],
                save_path = here::here("outputs", "figures", "BT_het_support.png"))
```

## Child Nutrition Index

```{r fig-child}
#| fig-height: 10
#| fig-width: 14
#| fig-cap: "Treatment effect heterogeneity — Child Nutrition Index (CH + CF + BFI)"

make_four_panel(results[["idx_child"]],
                save_path = here::here("outputs", "figures", "BT_het_child.png"))
```

---

# Interpretation

```{r interpretation}
#| code-fold: false
#| output: asis

for (r in results) {
  blp_sig <- r$blp_p < 0.20
  rate_sig <- r$rate_p < 0.10

  cat(glue("\n\n::: {{.callout-{ifelse(blp_sig, 'important', 'note')}}}\n"))
  cat(glue("## {r$domain$label}\n\n"))
  cat(glue("**BLP test:** p = {round(r$blp_p, 4)} — "))
  if (r$blp_p < 0.05) {
    cat("significant heterogeneity detected at 5% level.\n\n")
  } else if (r$blp_p < 0.10) {
    cat("marginally significant heterogeneity (p < 0.10).\n\n")
  } else if (r$blp_p < 0.20) {
    cat("suggestive heterogeneity (p < 0.20).\n\n")
  } else {
    cat("no significant heterogeneity detected.\n\n")
  }

  cat(glue("**RATE:** {round(r$rate_est, 4)} (SE = {round(r$rate_se, 4)}, p = {round(r$rate_p, 4)}) — "))
  if (rate_sig && r$rate_est > 0) {
    cat("positive targeting gains: prioritizing predicted high-responders would improve efficiency.\n\n")
  } else if (rate_sig && r$rate_est < 0) {
    cat("negative targeting signal: the families predicted to benefit most are NOT actually responding more.\n\n")
  } else {
    cat("no significant targeting gains from personalization.\n\n")
  }

  cat(glue("**CATE SD:** {round(r$cate_sd, 4)} — "))
  if (r$cate_sd < 0.05) {
    cat("very compressed distribution, minimal predicted variation.\n")
  } else if (r$cate_sd < 0.15) {
    cat("moderate predicted variation.\n")
  } else {
    cat("substantial predicted variation.\n")
  }

  cat(":::\n")
}
```

---

# Technical Notes

**Index construction:** Anderson (2008) inverse-covariance-weighted indices. Each component standardized to control-group mean = 0, SD = 1. Weights derived from the inverse of the pairwise covariance matrix among control-group components.

**Causal forest:** `grf::causal_forest()` with 4,000 trees, auto-tuned hyperparameters, honest splitting, clustered by township (TID). 15 baseline covariates (BLVAR_FEED_EBF excluded due to 39% structural missingness among pregnancy enrollees).

**BLP test:** Best Linear Projection calibration test (Chernozhukov, Demirer, Duflo, Fernández-Val 2018). Tests whether units predicted to have higher CATEs actually experience larger treatment effects.

**RATE:** Rank-Average Treatment Effect (Yadlowsky, Fleming, Shah, Sabatti, Wager 2021). Quantifies the expected gain from targeting based on predicted CATEs, using out-of-bag predictions for honest evaluation.
